### 2018-03-09 paper/background.tex
Finish Shake, Bazel, add overview draft

\noindent
So far so good: if either \cmd{main.exe} or \cmd{README} is dirty, \Make will
recreate the release archive. But now imagine that we need to specify the files
that go into \cmd{release.tar} in a text file \cmd{release.txt}. The dependencies
of \cmd{release.tar} are no longer known statically: they \emph{depend on the
contents} of \cmd{release.txt}, which might not even exist before the build
starts, e.g. if it is produced by concatenating input files \cmd{release-bins.txt}
and \cmd{release-docs.txt}. Alas, makefiles cannot express such \emph{dynamic
dependencies}, requiring ad hoc workarounds such as \emph{build phases}, which
are problematic~\cite{hadrian}.

@@

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
"release.tar" %> \_ -> do
    need ["release.txt"]
    contents <- lines <$> readFile "release.txt"
    need contents
    system "tar" $ ["-cf", "result.tar"] ++ contents
\end{minted}
\vspace{1mm}

\noindent
We first declare the static dependency on \cmd{release.txt}, then read its
content and depend on it, dynamically. Finally, we specify the command to
produce the resulting archive. Crucially, the archive will only be rebuilt if
one of the dependencies (static or dynamic) has changed.

\todo{AM}{How is \Shake different.}

@@

...

@@

When build systems are used by large teams, different team members
often end up executing exactly the same tasks on their local machines.
A \emph{cloud build system} can speed up builds dramatically by
transparently sharing build results among team members. Furthermore, cloud
build systems allow one to perform \emph{shallow builds} that materialise
only end build products on a local machine, leaving all intermediates in the
cloud. This is a significant optimisation compared to \emph{deep builds}
that require all transitive dependencies of an end build product to be
locally available. Non-cloud build systems cannot support shallow builds.

>>>>

\noindent
So far so good: if either \cmd{main.exe} or \cmd{README} is dirty, \Make will
recreate the release archive. But now imagine that we need to specify the files
that go into \cmd{release.tar} in a text file \cmd{release.txt}. The dependencies
of \cmd{release.tar} are no longer known statically: they \emph{depend on the
content} of \cmd{release.txt}, which might not even exist before the build
starts, e.g. if it is produced by concatenating input files \cmd{release-bins.txt}
and \cmd{release-docs.txt}. Alas, makefiles cannot express such \emph{dynamic
dependencies}, requiring ad hoc workarounds such as \emph{build phases}, which
are problematic~\cite{hadrian}.

@@

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
"release.tar" %> \_ -> do
    need ["release.txt"]
    files <- lines <$> readFile "release.txt"
    need files
    system "tar" $ ["-cf", "result.tar"] ++ files
\end{minted}
\vspace{1mm}

\noindent
We first declare the static dependency on \cmd{release.txt}, then read its
content (a list of files) and depend on it, dynamically. Finally, we specify the
command to produce the resulting archive. Crucially, the archive will only be
rebuilt if one of the dependencies (static or dynamic) has changed.

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency graph produced in the previous build to detect
which files are dirty -- a powerful idea with a long history of use in
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting}~\cite{acar2007selfadjusting} computation (see a discussion
in related work~\S\ref{sec-related}). Second, instead of abandoning and deferring
the computation of tasks whose rediscovered dependencies are dirty (as \Excel
does), \Shake \emph{pauses} their computation until the dependencies are
brought up-to-date.

@@

@@

When build systems are used by large teams, different team members
often end up executing exactly the same tasks on their local machines.
A \emph{cloud build system} can speed up builds dramatically by
transparently sharing build results among team members. Furthermore, cloud
build systems allow one to perform \emph{shallow builds} that materialise
only end build products on a local machine, leaving all intermediates in the
cloud. This is a significant optimisation compared to \emph{deep builds}
that require all transitive dependencies of an end build product to be
locally available. % Non-cloud build systems cannot support shallow builds.

\Bazel is one of the first examples of openly available cloud build systems.
Like \Make, it does not support dynamic dependencies and can therefore benefit
from the simplicity of building tasks in the statically known topological order.
It is minimal and supports the early cutoff optimisation.

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed in~\S\ref{sec-engineering}.}.
The latter allows to bypass the execution of a task, by predicting the hash of
its result from the hashes of the dependencies, and subsequently fetch the file
from the cache. A concrete implementation is provided in~\S\ref{sec-examples}.

\subsection{Overview}
\label{sec-background-overview}

\begin{table}[t]
\small
\centering
\begin{tabular}{l||l|l|l|l||l|l}
Build system & Dependencies & Minimal & Early cutoff & Cloud & Metadata & Algorithm \\
\hline
\Make      & Static  & Yes & No  & No  & Dirty bit        & Topological   \\
\Excel     & Dynamic & No  & No  & No  & Dirty bit, chain & Chain + defer \\
\Shake     & Dynamic & Yes & Yes & No  & Graph            & Recursive     \\
\Bazel     & Static  & Yes & Yes & Yes & Graphs, cache    & Topological   \\
\hline
This paper & Dynamic & Yes & Yes & Yes & Graphs, cache    & Recursive     \\
\end{tabular}
\vspace{2mm}
\caption{Overview of build systems.\label{tab-overview}}
\label{tab:use-cases}
\end{table}

### 2018-03-09 paper/related.tex
Finish Shake, Bazel, add overview draft

\begin{itemize}
    \item Memoization
    \item Self-adjusting computation
    \item Lens-like formulation
\end{itemize}

>>>>

\begin{itemize}
    \item Memoization
    \item Self-adjusting computation
    \item Lens-like formulation
    \item Other build systems: \Ninja, \cmd{nix},
          cloud (Blaze -- first? \Buck, Pants, CloudBuild, XXX),
          Shake derivatives (Jane Street's \cmd{Jenga}, Apple's \cmd{swift-llbuild}).
\end{itemize}

John Ericson suggested in a blog comment that nix may be somewhat monadic, see:
\url{https://blogs.ncl.ac.uk/andreymokhov/cloud-and-dynamic-builds/\#comment-1849}.
